---
title: "Linear Models"
author: "Sergio Ozoria"
date: "2025-11-06"
output: github_document
---

lm --> linear models
glm --> for generalized linear models. Make sure you're talking about the right distributions. Specify more stuff than a linear model

Use the broom package to help with complex linear regression output

```{r}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

Loading dataset

```{r}
data("nyc_airbnb")
```

Look at dataset / clean

```{r}
nyc_airbnb =
  nyc_airbnb |> 
  mutate(
    stars = review_scores_location / 2
  ) |> 
  rename(
    borough = neighbourhood_group
  ) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, room_type, neighbourhood)
```

Running linear regression

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb) # borough character variable and converts them into a factor by default and in alphabetical order
```

Additional cleaning

```{r}
nyc_airbnb =
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough), # useful for putting categories in how commonly they appear in the dataset
    room_type = fct_infreq(room_type)
  )

fit = lm(price ~ stars + borough, data = nyc_airbnb) 
```


```{r, eval = FALSE}
summary(fit)
names(summary(fit))
summary(fit)[["coefficients"]]
summary(fit)[["df"]]

fitted.values(fit)
resid(fit)
```


```{r}
fit |> 
  broom::tidy() |> 
  mutate(
    term = str_replace(term, "borough", "Borough: ")) |> 
  select(term, estimate, p.value) |> 
  knitr::kable(digits = 3)

fit |> 
  broom::glance()
```

Diagnostics(residuals)

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = resid, y = borough)) +
  geom_violin() 

nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = stars, y = resid)) +
  geom_point() 
```

Hypothesis testing

```{r}
fit |> 
  broom::tidy()
```

What about a categorical one

```{r}
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)

fit_alt |> 
  broom::tidy()

anova(fit_null, fit_alt) |> 
  broom::tidy()
```

Interactions vs nested data

```{r}
fit_interactions =
  lm(price ~ stars * borough + room_type * borough, data = nyc_airbnb)

fit_interactions |> 
  broom::tidy()
```


```{r}
nyc_airbnb |> 
  filter(borough == "Brooklyn") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()
```

Write a short function

```{r}
lm_airbnb = function(df) {
  
  lm(price ~ stars + room_type, data = df)
}

nyc_airbnb |> 
  filter(borough == "Brooklyn") |> 
  lm_airbnb() |> 
  broom::tidy()
```

